---
title: "Sentiment Analysis"
author: "Audrey Pentz"
date: "September 20, 2018"
output: html_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  library(tidyverse),
  library(tidytext),
  library(stringr),
  library(lubridate),
  library(knitr),
  library(qdap),            # polarity plots
  library(widyr),           # pairwise correlation
  library(ggrepel),         # geom_label_repel
  library(gridExtra),       # grid.arrange for multi-graphs
  library(kableExtra),      # tables
  library(formattable),     # color_tile
  library(yarrr),           # pirate plot
  library(radarchart),      # radar chart
  library(igraph),          # network diagrams
  library(ggraph),          # network diagrams
  library(wordcloud2)       # wordcloud
)

```


  
## **Load the Preprocessed Data**  
  
```{r import}

load("../input_data.RData")

```

  

## **Remove Stop Words and Tokenize by Sentences, Words and Bigrams**  
  
```{r tokenize}

# sentence tokenization
tidy_sona_sentences <- input_data$sentences

# word tokenization
tidy_sona_words <- input_data$words %>% 
  filter(!word %in% stop_words$word, str_detect(word, "[a-z]"))  # remove stop words

# bigram tokenization
bigrams <- tidy_sona_sentences %>% 
  unnest_tokens(bigram, sentence, token = "ngrams", n = 2)
# separate the bigrams 
bigrams_separated <- bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")
# remove stop words
bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
# join up the bigrams again
tidy_sona_bigrams <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")

# view data
View(tidy_sona_sentences)
View(tidy_sona_words)
View(tidy_sona_bigrams)

```
  

### **Sentiment Analysis using "bing" lexicon**  
  
The "bing" lexicon encodes words as either "positive" or "negative". However, not all words used in the SONA speeches are in the lexicon so we need to adjust for this.
  
#### **Sentiment per President**   
  
Let's understand how many "positive" and "negative" words are used by each president across all their SONA speeches and create a metric called "sentiment" which is simply the total number of positive words minus the total number of negative words. We then adjust for the total number of words used from the lexicon in the "sentiment_score" metric.  
  
```{r bing}

# average sentiment per president
avg_sentiment <- tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>%
  group_by(president) %>% 
  count(sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
         sentiment_score = round( (sentiment / (positive + negative) * 100), 2)) %>% 
  arrange(desc(sentiment_score))

# plot avg words per president
avg_sentiment %>%
  ungroup(sentiment_score, president) %>%
  mutate(sentiment_score = color_bar("lightblue")(sentiment_score)) %>%
  kable("html", escape = FALSE, align = "l", caption = "Sentiment Score per President") %>%
  kable_styling(bootstrap_options = 
                  c("striped", "condensed", "bordered"), 
                  full_width = FALSE)

```
   
**Insights:**   
  
Of the 3 presidents that have made more than 1 SONA speech, Zuma has the highest sentiment score, followed by Mbeki and then Mandela. Zuma's sentiment score is nearly double Mandela's. It's interesting that the current President, Ramaphosa, has the second highest sentiment score, not far behind Zuma and only slightly ahead of Mbeki.


  
#### **What are the 10 positive words most frequently used by each president?**
  
```{r}

positive_words <- tidy_sona_words %>%
  group_by(president) %>% 
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "positive") %>%
  count(word) %>%
  arrange(desc(n)) %>%
  group_by(president) %>% 
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(president, n) %>%
  mutate(row = row_number())

positive_words %>%
  ggplot(aes(x = row, n, fill = president)) +
    geom_col(show.legend = NULL) +
    labs(x = NULL, y = "Number of times positive word is used") + 
    ggtitle("Positive Words most used by each President") +
    facet_wrap(~president, 
               ncol = 3, nrow = 3, 
               scales = "free") +
    scale_x_continuous( 
      breaks = positive_words$row, 
      labels = positive_words$word) +
    coord_flip()

```
  
**Insights:**   
  
De Klerk's most used words were "freedom", "peaceful" and "support" and at least 2 of these 3 come up in all the president's most used words. Mandela's most used words also include "progress", "improve", "reconciliation" and "commitment" which are all words indicating repair and a move towards something better. Mbeki uses many of the same words but also introduces "empowerment" which is a word carried through by Zuma and Ramaphosa and "success" which is carried through by Zuma. These words suggest progress in the move towards repair or something better, first spoken about by Mandela. Ramaphosa also introduces the words "confidence", "effectively", "enhance" and "efficient", which are words commonly seen in a business context and have not shown up in any other SA president's top 10 most frequently used words in a SONA since 1994.  


  
#### **Which of the positive words most frequently used are common across presidents?**
  
```{r}

total_speeches <- tidy_sona_words %>% 
                 group_by(president) %>% 
                 summarise(total = n())

tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "positive") %>%
  group_by(president) %>%
  count(word, sort = TRUE) %>%
  left_join(total_speeches) %>% 
  mutate(freq = n/total) %>% 
  filter(rank(desc(freq)) <= 10) %>%
  ggplot(aes(reorder(word,freq), freq, fill = president)) + 
  geom_col() + 
  coord_flip() + 
  xlab("") +
  facet_grid(.~ president) +
  ggtitle("Positive Words most used across Presidents")

```
  
**Insights:**    
  
Common positive words across post 1994 presidents include: "freedom", "regard", "support", "improve" and "progress". Words introduced by Mandela and unique to his speeches are: "restructuring", "reconciliation", "committment", "contribution" and "succeed". Mbeki introduces the words "empowerment", "comprehensive", "integrated" and "improving" into the top words used and this is unique to his speeches. Zuma uses the words "success", "reform" and "pleased" frequently and other presidents do not. Ramaphosa introduces the words "significant", "productive", "confidence" and "effectively" which have not yet been seen in the any other SA president's top 10 most frequently used words in a SONA since 1994.  



#### **What are the 10 negative words most used by each president?**  

```{r negative}

negative_words <- tidy_sona_words %>%
  group_by(president) %>% 
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "negative") %>%
  count(word) %>%
  arrange(desc(n)) %>%
  group_by(president) %>% 
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(president, n) %>%
  mutate(row = row_number())

negative_words %>%
  ggplot(aes(x = row, n, fill = president)) +
    geom_col(show.legend = NULL) +
    labs(x = NULL, y = "Number of times negative word is used") + 
    ggtitle("Negative Words most used by each President") +
    facet_wrap(~president, 
               ncol = 3, nrow = 3, 
               scales = "free") +
    scale_x_continuous( 
      breaks = negative_words$row, 
      labels = negative_words$word) +
    coord_flip()

```
  
**Insights:**   
  
Common negative words pre 1994 include: "concerns"/"concern"/"concerned", "unconstitutional", "illusion", "hopeless", "disagree", "deprive", "conflict", and "boycott".

Common negative words post 1994 include: "corruption", "crime"/"criminal", "poverty"/"poor", "inequality", "issue"/"issues" and "crisis".

A negative word introduced by and unique to Mandela's top 10 is "struggle". 
Mbeki is the only president with the word "racism" in his top 10 negative words.
Motlanthe has "conflict" in his top 10 which no other president does. Zuma has "rail" which likely refers to the railway system and does have negative connotations for South Africa. Both Zuma and Ramaphosa use the word "difficult" a lot. Ramaphosa introduces the word "expropriation" into the top 10 for the first time amongst his peers.



#### **How many of the negative words most used were used by each president?**  
    
```{r}

tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "negative") %>%
  group_by(president) %>%
  count(word, sort = TRUE) %>%
  left_join(total_speeches) %>% 
  mutate(freq = n/total) %>% 
  filter(rank(desc(freq)) <= 10) %>%
  ggplot(aes(reorder(word,freq), freq, fill = president)) + 
  geom_col() + 
  coord_flip() + 
  xlab("") +
  facet_grid(.~ president) +
  ggtitle("Negative Words most used across Presidents")
  
```
  
**Insights:**   
  
The interpretation is much the same as before. Note the clear separation between the top 10 negative words used pre and post 1994 elections, indicative of the pre and post apartheid regimes.



#### **What proportion of words used are positive vs negative?**

```{r trends}

sentiments_per_year <- tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>%
  group_by(year, sentiment) %>%
  summarize(n = n()) 

sentiments_relative <- sentiments_per_year %>% 
  left_join(sentiments_per_year %>% 
            group_by(year) %>% 
            summarise(total = sum(n))) %>%
  mutate(freq = n/total) 

# plot proportion of positive and negative words used
ggplot(filter(sentiments_relative), aes(x = year, y = freq, fill = sentiment)) +
  geom_col() + coord_flip() + ggtitle("% of Negative and Positive Words per Year")

```
  
**Insights:**   
  
...



### **Change in Sentiment over time**

```{r}

year_total <- tidy_sona_words %>% 
  group_by(year) %>% 
  count() %>% 
  rename(total = n)

# per year
sentiments_per_year <- tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>%
  group_by(year) %>% 
  count(sentiment) %>%
  left_join(year_total) %>% 
  mutate(freq = n / total * 100)

sentiments_per_year %>%
  ggplot(aes(x = as.numeric(year), y = freq, colour = sentiment)) +
  geom_point() + 
  geom_smooth(method = "loess", aes(colour = sentiment))

```
  
**Insights:**    
  
...



#### **Distribution of "bing" Sentiment per President**

```{r}

# per year, per president
sentiments_per_year <- tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>%
  group_by(year, president) %>% 
  count(sentiment) %>%
  spread(key = sentiment, value = n) %>% 
  left_join(year_total) %>% 
  mutate(sentiment = positive - negative,
         avg_sentiment = sentiment / (positive + negative) * 100)

# plot of sentiment by president
ggplot(sentiments_per_year, aes(x = president, y = avg_sentiment, colour = president)) + 
  geom_boxplot() + # draw a boxplot for each president
  ggtitle("Boxplots of bing Sentiment per President")

```

**Insights:**    
  
...



#### **Change in "bing" Sentiment over time**

```{r}

# per year
sentiments_per_year <- tidy_sona_words %>%
  inner_join(get_sentiments("bing")) %>%
  group_by(year, president) %>% 
  count(sentiment) %>%
  spread(key = sentiment, value = n) %>% 
  left_join(year_total) %>% 
  mutate(sentiment = positive - negative,
         avg_sentiment = sentiment / (positive + negative) * 100)

sentiments_per_year %>%
  ggplot(aes(x = as.numeric(year), y = avg_sentiment, colour = president)) +
  geom_point() + 
  geom_smooth(method = "loess", aes(colour = sentiment)) +
  ggtitle("Change in bing Sentiment over time")

```
  
**Insights:**  
  
...



### **How can we be sure whether average sentiment is increasing over time?**

```{r}

model <- lm(avg_sentiment ~ as.numeric(year), data = sentiments_per_year)
summary(model)

```
  
**Insights:**   
  
...



### **Sentiment Analysis using "afinn" lexicon (scale from -5 negative to +5 positive)**

```{r afinn}

tidy_sona_words %>%
  inner_join(get_sentiments("afinn")) %>%
  count(score) %>% 
  mutate(weighted_score = score * n) %>% 
  ungroup(weighted_score) %>%
  mutate(weighted_score = color_bar("lightblue")(weighted_score),
         n = color_bar("lightblue")(n)) %>%
  kable("html", escape = FALSE, align = "r", caption = "afinn Sentiment") %>%
  kable_styling(bootstrap_options = 
                  c("striped", "condensed", "bordered"), 
                  full_width = FALSE)

```
  
**Insights:**  
  
...


#### **Distribution of "afinn" Sentiment per President**

```{r}

# per year, per president
sentiments_per_year <- tidy_sona_words %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(year, president) %>% 
  count(score) %>%
  mutate(score = ifelse(is.na(score), 0, score),
         total_words = sum(n),
         afinn_sentiment = sum(score * n),   # weighted score
         afinn_sentiment_perc = round((afinn_sentiment/total_words*100),2)) %>% 
         # afinn sentiment as a percentage of total words in the sentence
  spread(key = score, value = n, fill = 0)

# plot of sentiment by president
ggplot(sentiments_per_year, aes(x = president, y = afinn_sentiment_perc, colour = president)) +   geom_boxplot() + # draw a boxplot for each president
  ggtitle("Boxplots of afinn sentiment per President")

```

**Insights:**    
  
...



#### **Change in "afinn" Sentiment over time**

```{r}

sentiments_per_year %>%
  ggplot(aes(x = as.numeric(year), y = afinn_sentiment, colour = president)) +
  geom_point() + 
  geom_smooth(method = "loess", aes(colour = afinn_sentiment_perc)) +
  ggtitle("Change in afinn Sentiment over time")

```



### **Sentiment Analysis using "nrc" lexicon (infers emotion with certain words)**

```{r}

president_total <- tidy_sona_words %>% 
  group_by(president) %>% 
  count() %>% 
  rename(total = n)

# nrc sentiment per president
nrc_sentiment <- tidy_sona_words %>%
  inner_join(get_sentiments("nrc")) %>%
  group_by(president, sentiment) %>% 
  summarise(word_count = n()) %>%
  left_join(president_total) %>% 
  mutate(word_perc = round((word_count/total*100),2)) %>% 
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, word_count)) %>%
  ggplot(aes(sentiment, word_perc, fill = president)) +
  facet_wrap(~president, 
           ncol = 3, nrow = 3, 
           scales = "free") +
  geom_col() +
  guides(fill = FALSE) +                       # turn off the legend
  labs(x = NULL, y = "Relative Word Count") +
  scale_y_continuous(limits = c(0, 20)) +      # hard code the axis limit
  ggtitle("NRC Sentiment per President") +
  coord_flip()

```

**Insights:**   
...


### **Sentiment Analysis using "nrc" lexicon**

```{r}

nrc_sentiment <- tidy_sona_words %>%
  inner_join(get_sentiments("nrc")) %>%
  group_by(sentiment) %>% 
  count(word, sort = TRUE) %>%
  arrange(desc(n)) %>%
  slice(seq_len(8)) %>% #consider top_n() from dplyr also
  ungroup() 

nrc_sentiment %>% 
  # set `y = 1` to just plot one variable and use word as the label
  ggplot(aes(word, 1, label = word, fill = sentiment )) +
  # we want the words, not the points
  geom_point(color = "transparent") +
  # make sure the labels don't overlap
  geom_label_repel(force = 1,nudge_y = .5,  
                   direction = "y",
                   box.padding = 0.04,
                   segment.color = "transparent",
                   size = 3) +
  facet_grid(~sentiment) +
  theme(axis.text.y = element_blank(), axis.text.x = element_blank(),
        axis.title.x = element_text(size = 6),legend.position="none",
        panel.grid = element_blank(), panel.background = element_blank(),
        panel.border = element_rect("lightgray", fill = NA),
        strip.text.x = element_text(size = 9)) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("NRC Sentiment") +
  coord_flip()

```


### STILL TO DO:
### Bigrams
### Dealing with negation



### References

https://www.kaggle.com/rtatman/tutorial-sentiment-analysis-in-r

https://www.datacamp.com/community/tutorials/sentiment-analysis-R





